## Introduction

In this new analysis I will add two variables:  

 - Agriculture (% of GDP)  
 - Continent
 
The main goal of the analysis is to explore the relation between all the variables
across all the continents.
 
## Getting and Cleaning Data
Data has been obtained and downloaded from [Gapminder](http://www.gapminder.org/data/).

There are 3 datasets:  

 - **Agricultural land (% of land area)**,  
 - **Forest coverage (%)**,   
 - **Agriculture (% of GDP), **  
 - **Countries per Continent **  
 
 
**Note:** Since the goal of this course is not Getting and Cleaning data, I've
manually removed years from Agriculture data to make it easier to read and clean. 
*This is not a recomended practice*.

```{r,echo=FALSE, message=FALSE}
library(reshape2) # melt function
library(ggplot2)
library(gridExtra)
library(dplyr) # arrange
library(psych)
setwd("~/Repos//data-analysis-with-r")
```

```{r,warning=FALSE}
a.coverage <- read.csv("./data//agriculture land.csv",
                   col.names = c("Country","1990","2000","2005"))
f.coverage <- read.csv("./data//indicator_forest coverage.csv",
                   col.names = c("Country","1990","2000","2005"))
a.gdp <- read.csv("./data//agriculture p of GDP.csv", 
                  col.names = c("Country","1990","2000","2005"))
continents <- read.csv("./data//Countries-Continents.csv")

a.coverage <- melt(a.coverage, id.vars = c("Country"))
names(a.coverage) <- c("Country","Year","Agriculture.Coverage")
f.coverage <- melt(f.coverage, id.vars = c("Country"))
names(f.coverage) <- c("Country","Year","Forest.Coverage")
a.gdp <- melt(a.gdp, id.vars = c("Country"))
names(a.gdp) <- c("Country","Year","Agriculture.GDP")


data <- merge(a.coverage,f.coverage, by=c("Country","Year"))
data <- merge(data, a.gdp, by =c("Country", "Year"))
data <- merge(data, continents, by = c("Country"))

data$Year <- gsub("X","",data$Year)
data$Year <- as.factor(data$Year)
data$Agriculture.Coverage <- gsub(",",".",data$Agriculture.Coverage)
data$Agriculture.Coverage <- as.numeric(data$Agriculture.Coverage)
data$Forest.Coverage <- gsub(",",".",data$Forest.Coverage)
data$Forest.Coverage <- as.numeric(data$Forest.Coverage)
data$Agriculture.GDP <- gsub(",",".", data$Agriculture.GDP)
data$Agriculture.GDP <- as.numeric(data$Agriculture.GDP)

data <- data[complete.cases(data),]

rm(a.coverage,f.coverage, a.gdp, continents)

head(data)
```

Seems that data is ready to be processed...

## Data Processing

Let's do some explorative charts, nothing too formal, just to take a look at the
data...

```{r}
pairs.panels(data)
```

There isn't too much correlation between variables. The most interesting relation is
between Agriculture.Coverage and Forest.Coverage as I analized in my previuous exercise.
By looking at the Agriculture.GDP histogram it seems that only a few countries depends
on its agriculture.

Let's take a closer look to Agriculture.GDP in 2005:
```{r}
summary(subset(data, Year == 2005)$Agriculture.GDP)

ggplot(subset(data, Year == 2005), aes(x = Agriculture.GDP)) +
  geom_histogram(binwidth = 1)

ggplot(subset(data, Year == 2005), aes(x= Agriculture.GDP)) + 
  geom_histogram(aes(fill = Continent),binwidth = 1) + 
  facet_wrap(~ Continent)
```

#### Plotting Variable Means over Time
```{r}
ggplot(data, aes(x = Year, y= Agriculture.Coverage, group=Continent)) +
  geom_line(stat = "summary", aes(color = Continent), fun.y = mean) +
  ggtitle("Agriculture Coverage mean per Continent")

ggplot(data, aes(x = Year, y= Forest.Coverage, group=Continent)) +
  geom_line(stat = "summary", aes(color = Continent), fun.y = mean) +
  ggtitle("Forest Coverage mean per Continent")

ggplot(data, aes(x = Year, y= Agriculture.GDP, group=Continent)) +
  geom_line(stat = "summary", aes(color = Continent), fun.y = mean) +
  ggtitle("Agriculture % of GDP Coverage mean per Continent")
```

It looks like all the Agriculture % of GDP decreased in the World from 1990 to
2005 **even when** Agriculture Coverage increased in Africa, Asia, North America, and
South America.


#### Agriculture Ratio

Can **Agriculture.GDP** over **Agriculture.Coverage** may be used as a measure for the
performance of the farms? 

If two countries have different Agriculture.GDP but the same
Agriculture.Coverage could imply that the biggest Agriculture.GDP has a better use of
the soil. If 10% of my land produces the 20% of my GDP, then I'm making a better use
of the soil than a country that uses 10% of their land to produce the 5% of the GDP.

```{r}
data$Agriculture.Ratio <- data$Agriculture.GDP / data$Agriculture.Coverage
summary(data$Agriculture.Ratio)

ggplot(subset(data, Agriculture.Ratio < quantile(Agriculture.Ratio, probs = 0.95)),
              aes(x = Year, y = Agriculture.Ratio)) + geom_boxplot()

ggplot(subset(data, Agriculture.Ratio < quantile(Agriculture.Ratio, probs = 0.95)),
              aes(x = Year, y = Agriculture.Ratio)) + geom_boxplot() + facet_wrap(~Continent)
```

## Conclusion

Facts extracted so far:  

 - It seems that around the world, the Forest Coverage % per Country didn't change so much,
 - This doesn't mean that the Total Forest Coverage in the world didn't change drastically, but this can mean that most of the countries has a firm policy about preserving forest.
 - This also can mean that the main deforestation in the world is being done by a few countries.
 - Even when countries are spending more and more land on Agriculture this doesn't mean that its contribution to the GDP is increasing, just the oposite.
 - The most remarcable case is Asia, that has increased its Agriculture Coverage by almost 10% but the Agriculture.GDP decreased by 3%

```{r}
ggplot(subset(data, Continent == "Asia"),
       aes(x = Year, y = Agriculture.Coverage, group = Continent)) +
  geom_line(stat = "summary", aes(color = Continent), fun.y = mean)

ggplot(subset(data, Continent == "Asia"),
       aes(x = Year, y = Agriculture.GDP, group = Continent)) +
  geom_line(stat = "summary", aes(color = Continent), fun.y = mean)


```



