# Introduction
This analysis is the result of an excercise made in Udacity's Course: [Exploratory Data Analysis with R](https://www.udacity.com/course/ud651)

### Excercise Instructions: Gapminder Data
The Gapminder website contains over 500 data sets with information about the world's population. Your task is to download a data set of your choice and create 2-5 
plots that make use of the techniques from Lesson 3.

You might use a simple histogram, a boxplot split over a categorical variable,
or a frequency polygon. The choice is yours!

[Link to the Gapminder website](http://www.gapminder.org/data/).

Once you've completed your investigation, create a post in the discussions that includes:  
1. any questions you answered, your observations, and summary statistics  
2. snippets of code that created the plots  
3. links to the images of your plots  

# Data Analysis  
#### Load necesary Libraries
```{r}
library(reshape2)
library(plyr)
library(ggplot2)
library(rworldmap)
library(RColorBrewer)
```


### Download and Cleaning data
```{r}
setwd("~/Repos/data-analysis-with-r/")
file.path <- "./data/indicator_forest coverage.csv"
data <- read.csv(file.path,col.names = c("Country","1990","2000","2005"))

data <- melt(data, id.vars = c("Country"))
names(data) <- c("Country","Year","Forest.Coverage")

data$Year <- gsub("X","",data$Year)
data$Year <- as.factor(data$Year)

data$Forest.Coverage <- gsub(",",".",data$Forest.Coverage)
data$Forest.Coverage <- as.numeric(data$Forest.Coverage,na.action)

data <- data[complete.cases(data),]
```
### Data Processing
```{r}
ggplot(data) + 
  geom_boxplot(aes(x = Year, y = Forest.Coverage, fill = Year), alpha = 0.5) + 
  scale_y_continuous(breaks = seq(0,60,5)) +
  coord_cartesian(ylim = c(10,60)) + 
  ylab("Average Forest Coverage (%)") + 
  ggtitle("Average Forest Coverage (%) by Year")
```
  
We can deduce that mean of the percent of Forest Coverage accross all the 
Countries decreased from ~30% to ~28% between 1990 and 2005.

#### Argentina:
```{r}
argentina <- subset(data, Country == "Argentina")
argentina

ggplot(argentina) + 
  geom_line(aes(x=Year,y=Forest.Coverage, group = Country), stat= "identity",
            size = 1, color = "red") +
  geom_boxplot(data = data , aes(x=Year,y=Forest.Coverage, fill = Year), alpha = 0.25) + 
  ggtitle("Forest Coverage (%) Changes in Argentina between 1990 and 2005\n 
          compared with the average of the World") + 
  ylab("Average Forest Coverage (%)") 
```
  
Argentina's Forest Coverage went down from 12.88% to 12.07%. Argentina's values
are limiting the First Quantile of the entire population.


#### South America
```{r}
south.america <- c("Argentina", "Chile", "Uruguay", "Paraguay", "Brazil",
                   "Peru", "Ecuador", "Bolivia", "Venezuela", "Colombia",
                   "Suriname", "French Guiana", "Guyana")
south.america.data <- subset(data, Country %in% south.america)

ggplot(south.america.data) + 
  geom_line(aes(x=Year,y=Forest.Coverage, group = Country, 
                color = Country),
            stat= "identity", size = 1) +
  ggtitle("Forest Coverage (%) Changes in South America Countries\n 
           between 1990 and 2005 compared with the average of the region.") + 
  geom_boxplot(aes(x=Year,y=Forest.Coverage, fill = Year), alpha = 0.25) + 
  ylab("Forest Coverage (%)") +  
  #scale_colour_manual(values = c("red","blue", "green", "yellow", "cyan", "grey",
   #                              "brown", "purple", "pink", "gold")) + 
  scale_y_continuous(breaks = seq(0,70,5))
```
  
Some conclusions:  
 - There are a lot of differences between the top countries, Brazil and Colombia,
 and Uruguay and Chile. Almost a 50% of difference.  
 - Only 2 countries has increased it's Forest Coverage, Uruguay and Chile.  
 - The mean of Forest Coverage in South America decreased from ~58% to ~50%. This 
 difference is bigger than the average of the World.  
 - Countries like Brazil, Colombia and Venezuela has almost a 60% of Forest Coverage!  

#### G20 Countries 
I've excluded European Union and South Korea due to missing values in the Data Set
```{r}
g20 <- c("Argentina","Australia", "Brazil","Canada","China","France",
                   "Germany", "India", "Indonesia","Italy", "Japan","Mexico",
                   "Russia", "Saudi Arabia", "South Africa",
                   "Turkey", "United Kingdom", "United States")
g20.data <- subset(data, Country %in% g20)

ggplot(g20.data) + 
  ggtitle("Forest Coverage (%) Changes in g20 Countries.\n 
           between 1990 and 2005,\n compared with the average of the group.") + 
  geom_boxplot(aes(x=Year,y=Forest.Coverage, fill = Year), alpha = 0.25) + 
  ylab("Forest Coverage (%)") +
  scale_y_continuous(breaks = seq(0,70,5))
```
  
We can deduce that mean of the Forest Coverage accross all the g20
Countries increased from ~27.5% to ~30% between 1990 and 2005. Since the World
tendency is to decrease, **Which Countries in the G20 Group have increased it's Forest Coverage?**

#### G20 Countries that increased its Forest Coverage
Now I want to calculate the variance in the Forest Coverage in the G20 Countries
to determine which ones increased and which ones decreased.  

```{r}
# Calculate de Difference Between FC in 2005 and FC in 1990.
Changes <- subset(g20.data, Year == 2005)$Forest.Coverage - subset(g20.data, Year == 1990)$Forest.Coverage
# Get the list of Countries
Country <- subset(g20.data, Year == 2005)$Country
# Create the Data Frame
g20.Changes <- data.frame(Country, Changes)
# Order the Data Frame by Changes
g20.Changes <- arrange(g20.Changes,Changes,decreasing = TRUE)
# Reset the Factors levels so ggplot can plot in correct order
g20.Changes$Country <- factor(g20.Changes$Country,
                                      levels = g20.Changes$Country) 
g20.Changes

ggplot(g20.Changes) + 
  geom_bar(aes(x=Country,y=Changes), fill = "blue4", stat = "identity") +
  scale_y_continuous(breaks = seq(-20,15,1)) + 
  ggtitle("Changes in Forest Coverage (%)\n in g20 countries from 1990 to 2005") +
  ylab("Changes in Forest Coverage (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```
  
Some interesting results:  
- Italy, China, France, India, United Kingdom, Germany, Turkey and United States
 **increased** their Forest Coverage from 1990 to 2005  
- Japan, Australia, Argentina, Mexico, Brazil, and Indonesia **decreased**
 their Forest Coverage between 1900 and 2005  
- Wondering about Indonesia? [Deforestation in Indonesia](http://en.wikipedia.org/wiki/Deforestation_in_Indonesia)

#### Top Countries in the World 2005
Which are the Countries in the World with most Forest Coverage?  

```{r}
top.countries.2005 <- subset(data, Year == 2005)
top.countries.2005 <- top.countries.2005[order(top.countries.2005$Forest.Coverage, 
                                               decreasing = TRUE),]

top.countries.2005$Country <- factor(top.countries.2005$Country, 
                                     levels = top.countries.2005$Country)
# Plot Top 20
ggplot(top.countries.2005[1:20,]) + 
  geom_bar(aes(x=Country, y = Forest.Coverage), fill = "blue4", stat = "identity") +
  ylab("Forest Coverage in 2005") + 
  ggtitle("20 Countries with more Forest Coverage (%) in the World in 2005.") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
  
- Learn a little bit more about [Pitcairng](http://en.wikipedia.org/wiki/Pitcairn_Islands)  
- Some info about [Suriname](http://en.wikipedia.org/wiki/Suriname)

Which are the Countries in the World with less Forest Coverage?  

```{r}
# Plot Top 20
ggplot(tail(top.countries.2005, 20)) + 
  geom_bar(aes(x=Country, y = Forest.Coverage), fill = "blue4", stat = "identity") +
  ylab("Forest Coverage in 2005") + 
  ggtitle("20 Countries with less Forest Coverage (%) in the World in 2005.") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#### Top/Bottom Changes in Forest Coverage in the World
Which countries has increased/decreased its Forest Coverage between 1990 and 2005?  

```{r}
Country <- as.character(unique(data$Country))
Changes <- subset(data, Year == 2005)$Forest.Coverage - subset(data, Year == 1990)$Forest.Coverage

countries.Changes <- data.frame(Country, Changes)
countries.Changes <- arrange(countries.Changes,Changes,decreasing = T)
countries.Changes$Country <- factor(countries.Changes$Country,
                                      levels = countries.Changes$Country) 

ggplot(countries.Changes[1:20,]) + 
  geom_bar(aes(x=Country,y=Changes), fill = "blue4", stat = "identity") +
  scale_y_continuous(breaks = seq(-11,15,1)) + 
  ggtitle("Changes in Forest Coverage (%)\n from 1990 to 2005") +
  ylab("Changes in Forest Coverage (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(tail(countries.Changes, 20)) + 
  geom_bar(aes(x=Country,y=Changes), fill = "blue4", stat = "identity") +
  scale_y_continuous(breaks = seq(-25,15,1)) + 
  ggtitle("Changes in Forest Coverage (%)\n from 1990 to 2005") +
  ylab("Changes in Forest Coverage (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

x <- rbind(countries.Changes[1:10,], tail(countries.Changes, 10))
ggplot(x) + 
  geom_bar(aes(x=Country,y=Changes), fill = "blue4", stat = "identity") +
  scale_y_continuous(breaks = seq(-25,15,1)) + 
  ggtitle("Top 10 and Bottom 10 changes in Forest Coverage (%)\n from 1990 to 2005") +
  ylab("Changes in Forest Coverage (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
  
- Wow! Honduras has reduced its Forest Coverage by almost a 25%!


#### Fancy World Map Graph
Of course there should be a Fancy World Map with some info:
```{r}
colourPalette <- brewer.pal(7,"Greens")
sPDF <- joinCountryData2Map( countries.Changes, joinCode = "NAME", 
                             nameJoinColumn = "Country")
mapCountryData( sPDF, nameColumnToPlot="Changes",
                   colourPalette = colourPalette )
```
