---
title: "Forest Coverage vs Agriculture Coverage"
author: "pdelboca"
date: "30/01/2015"
output: html_document
---
## Introduction
In my last analysis I went through the Forest Coverage in the main Countries of 
the world. As now, I'm practicing 2 Variables analysis I will add the Agriculture 
Coverage of each country to analyze the relationship between deforestation and 
agriculture in each country.

```{r}
library(reshape2) # melt function
library(ggplot2)
library(gridExtra)
setwd("~/Repos//data-analysis-with-r")
```

## Getting and Cleaning Data
Data has been obtained and downloaded from [Gapminder](http://www.gapminder.org/data/).

There are 2 datasets: **Agricultural land (% of land area)** and **Forest coverage (%)**.

**Note:** Since the goal of this course is not Getting and Cleaning data, I've
manually removed years from Agriculture data to make it easier to read and clean. 
*This is not a recomended practice*.


```{r,warning=FALSE}
a.coverage <- read.csv("./data//agriculture land.csv",
                   col.names = c("Country","1990","2000","2005"))
f.coverage <- read.csv("./data//indicator_forest coverage.csv",
                   col.names = c("Country","1990","2000","2005"))

a.coverage <- melt(a.coverage, id.vars = c("Country"))
names(a.coverage) <- c("Country","Year","Agriculture.Coverage")
f.coverage <- melt(f.coverage, id.vars = c("Country"))
names(f.coverage) <- c("Country","Year","Forest.Coverage")

data <- merge(a.coverage,f.coverage, by=c("Country","Year"))

data$Year <- gsub("X","",data$Year)
data$Year <- as.factor(data$Year)
data$Agriculture.Coverage <- gsub(",",".",data$Agriculture.Coverage)
data$Agriculture.Coverage <- as.numeric(data$Agriculture.Coverage,na.action)
data$Forest.Coverage <- gsub(",",".",data$Forest.Coverage)
data$Forest.Coverage <- as.numeric(data$Forest.Coverage,na.action)

rm(a.coverage,f.coverage)

head(data)
```

Wow! I'm already learning something. I didn't know that [Afghanistan main economy
was Agriculture](http://en.wikipedia.org/wiki/Afghanistan#Economy)

## Data Processing

#### First some basic histograms of the Variables:

```{r}
p1 <- ggplot(data, aes(x=Forest.Coverage)) + geom_histogram(binwidth = 3) + 
  facet_wrap(~ Year) +
  ggtitle("Forest Coverage around the World per Year")

p2 <- ggplot(data, aes(x=Agriculture.Coverage)) + geom_histogram(binwidth = 3) + 
  facet_wrap(~ Year) +
  ggtitle("Agriculture Coverage around the World per Year")

grid.arrange(p1,p2,ncol = 1)
```

It look's like the Agriculture Coverage tends to a normal distribution. It's a 
pity I don't have data from the last 10 years!

#### Let's plot a few BoxPlots:

```{r, warning=FALSE}
ggplot(data, aes(x = Year, y = Forest.Coverage, fill = Year)) + 
  geom_boxplot(alpha = 0.5) + 
  scale_y_continuous(breaks = seq(0,100,5)) +
  coord_cartesian(ylim = c(00,100)) + 
  ylab("Average Forest Coverage (%)") + 
  ggtitle("Average Forest Coverage (%) by Year")


ggplot(data, aes(x = Year, y = Agriculture.Coverage, fill = Year)) + 
  geom_boxplot(alpha = 0.5) + 
  scale_y_continuous(breaks = seq(0,100,5)) +
  coord_cartesian(ylim = c(00,100)) + 
  ylab("Average Agriculture Coverage (%)") + 
  ggtitle("Average Agriculture Coverage (%) by Year")

```

Regarding Agriculture seems that 75% of the countries has 20% or more Agriculture 
Coverage and the median is almost 40%.

#### Argentina's Agriculture:

```{r}
argentina <- subset(data, Country == "Argentina")
argentina
```
Agriculture in Argentina has grown almost a 3% Between 1990 and 2005.

```{r}
ggplot(argentina) + 
  geom_line(aes(x=Year,y=Agriculture.Coverage, group = Country), stat= "identity",
            size = 1, color = "red") +
  geom_boxplot(data = data , aes(x=Year,y=Agriculture.Coverage, fill = Year), alpha = 0.25) + 
  ggtitle("Agriculture Coverage (%) Changes in Argentina between 1990 and 2005\n 
          compared with the average of the World") + 
  ylab("Average Agriculture Coverage (%)") 
```

It's not surprising to see Argentina closer to the 75% Quantile than the median, 
[Argentina has a long agriculture tradition](http://en.wikipedia.org/wiki/Agriculture_in_Argentina). Also remind that Argentina is the [8th biggest country in the world](http://simple.wikipedia.org/wiki/List_of_countries_by_area) which gives Argentina a lot of squared kilometers dedicated
to the agriculture.

#### Agriculture vs Forest Coverage

Now let's start plotting some scatterplots:

```{r}
ggplot(data, aes(x=Forest.Coverage, y=Agriculture.Coverage)) + geom_point()
```

.... Okay.... No idea of how to read this plot. Let's try plotting only 1 Year, 
lets say, the last one:

```{r}
ggplot(subset(data, Year == 2005),
       aes(x=Forest.Coverage, y=Agriculture.Coverage)) + 
  geom_point() +
  ggtitle("Forest Coverage vs Agriculture Coverage (2005)")
```


Still no idea... :_)

But I can see **a few interesting things**:  
* There is a Diagonal tendence with negative slope  
* It seems to be a Vertical line near the 0% of Forest Coverage  
* There are only a few points above the diagonal  

Regarding the third point... There shouldn't be points above the main diagonal,
that will mean that Agriculture Coverage + Forest Coverage is bigger than the 100%
of the land. Two options: **bad data** or some countries uses their forest as
agriculture land... Let's dig a little bit.

```{r}
ggplot(subset(data, Year == 2005),
       aes(x=Forest.Coverage, y=Agriculture.Coverage)) + 
  geom_point() +
  ggtitle("Forest Coverage vs Agriculture Coverage (2005)") + 
  geom_abline(intercept = 100, slope = -1) +
  scale_y_continuous(breaks = seq(0,100,5)) +
  scale_x_continuous(breaks = seq(0,100,5)) 

subset(data, Forest.Coverage + Agriculture.Coverage > 100)
```


It's logical to say that Forest Coverage and Agriculture Coverage has a negative
 correlation, after all, you need to deforest in order to have more land to grow crops.
 Let's see if there is a Correlation between the Forest Coverage and Agriculture Coverage
 for 2005:
 
 
```{r}
with(subset(data, Year == 2005),
     cor.test(Forest.Coverage,Agriculture.Coverage))
```


#### South America Countries
```{r}
south.america <- c("Argentina", "Chile", "Uruguay", "Paraguay", "Brazil",
                   "Peru", "Ecuador", "Bolivia", "Venezuela", "Colombia",
                   "Suriname", "French Guiana", "Guyana")
south.america.data <- subset(data, Country %in% south.america)

ggplot(subset(south.america.data, Year == 2005),
       aes(x=Forest.Coverage, y=Agriculture.Coverage)) + 
  geom_point(aes(color = Country), size = 5, alpha = 1/2) +
  ggtitle("Forest Coverage vs Agriculture Coverage (2005)")
  

with(subset(south.america.data, Year == 2005),
     cor.test(Forest.Coverage,Agriculture.Coverage))

```


#### G20 Countries

