Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
setwd("~/Repos//data-analysis-with-r")
library(ggplot2)
pf <- read.csv("./data/pseudo_facebook.tsv", sep ="\t")
p <- ggplot(pf) + theme_bw()
p + geom_point(aes(x=age, y=friend_count))
```

***

#### What are some things that you notice right away?
Response:
A lot of outliers. A lot of people with age 90+, also the majority of the friends are under the age of 30.
***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
p + geom_jitter(aes(x=age, y=friend_count), alpha = 1/20) +
  xlim(13,90)
``

### Coord_trans()
Notes:

```{r Coord_trans()}
p + geom_point(aes(x=age, y=friend_count), alpha = 1/20) +
  xlim(13,90) + coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Explore the relation between friends initiated vs age
Notes:

```{r Alpha and Jitter}
p + geom_jitter(aes(x=age,y=friendships_initiated), alpha = 1/20) +
  xlim(13,90) 
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

age_group <- group_by(age, .data = pf)
pf.fc_by_age <- summarise(age_group,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n()
                          )
pf.fc_by_age <- arrange(pf.fc_by_age, age)


ggplot(pf.fc_by_age) + geom_line(aes(x=age, y=friend_count_mean), color = "red") +
  ggtitle("Friend Count Mean/Median by Age") +
  xlim(13,90)

ggplot(pf.fc_by_age) + geom_line(aes(x=age, y=friend_count_median), color = "blue") +
  ggtitle("Friend Count Median by Age") +
  xlim(13,90)

```
***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

