Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
setwd("~/Repos//data-analysis-with-r")
library(ggplot2)
pf <- read.csv("./data/pseudo_facebook.tsv", sep ="\t")
p <- ggplot(pf) + theme_bw()
p + geom_point(aes(x=age, y=friend_count))
```

***

#### What are some things that you notice right away?
Response:
A lot of outliers. A lot of people with age 90+, also the majority of the friends are under the age of 30.
***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
p + geom_jitter(aes(x=age, y=friend_count), alpha = 1/20) +
  xlim(13,90)
``

### Coord_trans()
Notes:

```{r Coord_trans()}
p + geom_point(aes(x=age, y=friend_count), alpha = 1/20) +
  xlim(13,90) + coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Explore the relation between friends initiated vs age
Notes:

```{r Alpha and Jitter}
p + geom_jitter(aes(x=age,y=friendships_initiated), alpha = 1/20) +
  xlim(13,90) 
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

age_group <- group_by(age, .data = pf)
pf.fc_by_age <- summarise(age_group,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n()
                          )
pf.fc_by_age <- arrange(pf.fc_by_age, age)


ggplot(pf.fc_by_age) + geom_line(aes(x=age, y=friend_count_mean), color = "red") +
  ggtitle("Friend Count Mean/Median by Age") +
  xlim(13,90)

ggplot(pf.fc_by_age) + geom_line(aes(x=age, y=friend_count_median), color = "blue") +
  ggtitle("Friend Count Median by Age") +
  xlim(13,90)

```
***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age,y=friend_count), data = pf) + geom_point(alpha = 1/20, color = "orange") +
  geom_line(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = quantile, prob = 0.1, linetype = 2, color = "red") + 
  geom_line(stat = "summary", fun.y = quantile, prob = 0.5, color = "red") + 
  geom_line(stat = "summary", fun.y = quantile, prob = 0.9, linetype = 2, color = "red") + 
  coord_cartesian(xlim = c(30,60)) + 
  coord_trans(y = "sqrt")
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count)
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count, method = "pearson"))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(pf) + geom_point(aes(x=www_likes_received,y=likes_received), color = "blue")
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(pf,aes(x=www_likes_received,y=likes_received)) + geom_point(color = "blue") +
  xlim(0, quantile(pf$www_likes_received,prob=0.95)) +
  ylim(0, quantile(pf$likes_received,prob=0.95)) + 
  geom_smooth(method = "lm",color = "red", size = 2)

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received,pf$likes_received)
```

Response: 0.948
***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}
?Mitchell
ggplot(Mitchell) + geom_point(aes(x=Month,y=Temp))
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0.1
b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor(Mitchell$Month,Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point() +
  scale_x_discrete(breaks=seq(0, 203, 12), labels=(seq(1976,1992,1))) +
xlab('Year')

ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point()
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(data = pf.fc_by_age, aes(x=age,y=friend_count_median)) + geom_line()
```

***

### Age with Months Means

```{r Age with Months Means}
str(pf)
pf$age_with_months <- pf$age + (12 - pf$dob_month)/12
```

Programming Assignment
```{r Programming Assignment}
library(dplyr)
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(subset(pf.fc_by_age_months, age_with_months < 71),
      aes(x = age_with_months, y = friend_count_mean)) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

