Problem Set 3: Diamonds
=======================
This is a file to solve all the excercises of **Problem Set 3: Diamonds** 
from Udacity's Course: Data Analysis with R.

### Frist Excercise:
Basic exploration of Diamonds data set.
```{r}
library(ggplot2)
data(diamonds)

dim(diamonds) # 53940 obs and 10 variables
str(diamonds) # 3 Ordered Factors
str(diamonds$color)
?diamonds # D Represents the best Color
```

### Second Excercise:
Create a histogram of the price of all the diamonds in the diamond data set.
```{r}
ggplot(diamonds) + geom_histogram(aes(x=price), binwidth = 100, 
                                  color = "black", fill = "lightblue") + 
  ggtitle("Diamonds Price Histogram")
```

### Excercise 3:
Describe the shape and center of the price distribution. Include summary 
statistics like the mean and median
```{r}
summary(diamonds$price)
```
The shape of the distribution is **Skewed Right** with its **center in 2401**. I chose 
the Median as center due to the long tail of the graph. The mean of the distribution
is 3933.

### Excercise 4: Diamond Counts
How many diamonds cost less than $500?
```{r}
nrow(subset(diamonds, price < 500))
```
How many diamonds cost less than $250?
```{r}
nrow(subset(diamonds, price < 250))
```
How many diamonds cost $15000 or more?
```{r}
nrow(subset(diamonds, price >= 15000))

```

### Excercise 5: Exploring the Peak

```{r}
ggplot(diamonds) + geom_histogram(aes(x=price), binwidth = 10, 
                                  color = "black", fill = "lightblue") + 
  ggtitle("Diamonds Price Histogram between $0 and $1500.") + 
  coord_cartesian(xlim=c(0,1500))
```

### Excercise 6: Price by cut histogram
Break out the histogram of diamond prices by cut. You should have five histograms 
in separate panels on your resulting plot.  
```{r}
ggplot(diamonds) + geom_histogram(aes(x=price), binwidth = 100, 
                                  color = "black", fill = "lightblue") + 
  ggtitle("Diamonds Price by Cut Histogram") + 
  facet_grid(. ~ cut)
```
   
### Excercise 7: Price by Cut
Answer the question below. Check more than one option if there are ties.

Which cut has the highest priced diamond?
```{r}
diamonds[which.max(diamonds$price),]
#or
summary(diamonds$price) # Max = 18820
diamonds[diamonds$price == 18823,] # 18820 doesn't exist!?!?!?

```
Which cut has the lowest priced diamond?
```{r}
which.min(diamonds$price)
# or
summary(diamonds$price) # Min = 326
diamonds[diamonds$price == 326,]
```
Which cut has the lowest median price?
```{r}
by(diamonds$price,diamonds$cut,summary)
```

### Excercise 8: Scales and Multiple Histograms
```{r}
ggplot(diamonds) + geom_histogram(aes(x=price), binwidth = 100, 
                                  color = "black", fill = "lightblue") + 
  ggtitle("Diamonds Price by Cut Histogram") + 
  facet_grid(. ~ cut)

ggplot(diamonds) + geom_histogram(aes(x=price)) + 
  ggtitle("Diamonds Price by Cut Histogram") + 
  facet_wrap(~ cut,scales="free_y",ncol = 5)

qplot(x = price, data = diamonds) + facet_wrap(~cut, scales="free")
```

### Excercise 9: Price per Carat by Cut
Create a histogram of price per carat  and facet it by cut. You can make adjustments
to the code from the previous exercise to get started.

Adjust the bin width and transform the scale of the x-axis using log10.

```{r}
ggplot(diamonds) + geom_histogram(aes(x=price/carat), binwidth = 0.05,
                                  color = "black", fill = "lightblue") +
  ggtitle("Histogram of Price per Carat, facet by Cut.") + 
  scale_x_log10() +
  facet_grid(. ~ cut)
```

### Excercise 10: Price Box Plots
Investigate the price of diamonds using box plots, numerical summaries, and one 
of the following categorical variables: cut, clarity, or color.

```{r}
by(diamonds$price, diamonds$color, summary)
ggplot(diamonds) + geom_boxplot(aes(x = color,  y = price, fill = color)) + 
  coord_cartesian(ylim = c(0, 8000)) + 
  scale_y_continuous(breaks = seq(0,8000,500)) +
  ggtitle("Diamonds Price by Color.")
```

### Excercise 11: Interquartile range - IQR
a) What is the price range for the middle 50% of diamonds with color D?
```{r}
by(diamonds$price, diamonds$color, summary)
```
b) What is the price range for the middle 50% of diamonds with color J?
```{r}
by(diamonds$price, diamonds$color, summary)
```
c) What is the IQR for diamonds  with the best color?
```{r}
by(diamonds$price, diamonds$color, IQR)
```

d) What is the IQR for diamonds  with the worst color?
```{r}
by(diamonds$price, diamonds$color, IQR)
```

### Excercise 12: Price per Carat Box Plots by Color
Investigate the price per carat of diamonds across the different colors of diamonds 
using boxplots.
```{r}
ggplot(diamonds) + geom_boxplot(aes(x=color,y=price/carat, fill=color)) + 
  scale_y_continuous(breaks = seq(0,8000,500)) + 
  coord_cartesian(ylim = c(0,8000)) + 
  ylab("Price per Carat") +
  ggtitle("Price per Carat by Color")
```

### Carat frequency polygon
Investigate the weight of the diamonds (carat) using a frequency polygon. Use different
bin widths to see how the frequency polygon changes. What carat size has a count greater 
than 2000?
```{r}
ggplot(diamonds) + geom_freqpoly(aes(x = carat), binwidth = 0.02) + 
  scale_x_continuous(breaks = seq(0,5,0.1)) + 
  geom_hline(y = 2000,color = "red") + 
  ggtitle("Carat Frequency Polygon")
```

### Excercise 13: Gapminder data

[Link to the Excercise](http://rpubs.com/pdelboca/forest-coverage)

### Excercise 14: Birthday Histograms
Your task is to investigate the distribution of your friends' birth months and days.  
Here some questions you could answer, and we hope you think of others.

**********************************************************************

- How many people share your birthday? Do you know them?  
- Which month contains the most number of birthdays?  
- How many birthdays are in each month?  
- Which day of the year has the most number of birthdays?  
- Do you have at least 365 friends that have birthdays on everyday
of the year?  

#### Load and Cleaning data
```{r}
library(lubridate)
setwd("~/Repos/data-analysis-with-r/")
birthdays <- read.csv("./data//birthdaysExample.csv")
birthdays$dates <- mdy(birthdays$dates)
birthdays$month <- month(birthdays$dates)
birthdays$day <- day(birthdays$dates)
```
#### Data Processing
Some basic analysis:
```{r}
summary(birthdays)
table(birthdays$month)
table(birthdays$day)
```


How many people share your birthday? Do you know them?
```{r}
# April, 3rd
april.third <- subset(birthdays, month == 04 & day == 03)
nrow(april.third)
```
Which month contains the most number of birthdays?
```{r}
ggplot(birthdays) + geom_histogram(aes(x = month),
                                   color = I('black'),
                                   fill = I('#099DD9'),
                                   binwidth = 1) +
  scale_y_discrete() +
  scale_x_discrete() + 
  ggtitle("Birthdays per Month")
```
How many birthdays are in each month?
```{r}
table(birthdays$month)
```
Which day of the year has the most number of birthdays? 
```{r}
ggplot(birthdays) + geom_histogram(aes(x = day),
                                   color = I('black'),
                                   fill = I('#099DD9'),
                                   binwidth = 1) +
  scale_y_discrete() +
  scale_x_discrete() + 
  ggtitle("Birthdays per Day")
```
Do you have at least 365 friends that have birthdays on everyday
of the year? 
```{r}
length(unique(birthdays$dates))
# There are 348 different dates so the answer is no
```





